A data-intensive assessment of the species-abundance distribution.

### 2015-04-03 defense_pres.tex
Moves open science slide.
Reorders, adds slides.

\documentclass[14pt]{beamer}
\usetheme{Pittsburgh}
\usepackage{xcolor}
\definecolor{USUBlue}{RGB}{26,57,89}
\usecolortheme[named=USUBlue]{structure}
\setbeamercolor{normal text}{fg=USUBlue}
\usepackage[utf8]{inputenc}
\usepackage{mathptmx}
\usepackage{tgbonum}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{marvosym}
\usefonttheme{structuresmallcapsserif}
\usefonttheme{serif}
\setbeamercolor{author}{fg=USUBlue}
\setbeamerfont{author}{size=\small}
\setbeamerfont{frametitle}{size=\large}
\setbeamertemplate{enumerate items}[default]
\author{Elita Baldridge}
\title[17pt]{A data-intensive assessment of the species-abundance distribution.}
\setbeamertemplate{navigation symbols}{}
\date{}
%\setbeamercovered{transparent} 
%\logo{\includegraphics[scale=.03]{../Miscellaneous/Pictures/ecology_center_horizontal.jpg}\includegraphics[scale=0.07]{../Miscellaneous/Pictures/Weecology.png}} 
\institute{\includegraphics[scale=.07]{../Miscellaneous/Pictures/ecology_center_horizontal.jpg}\includegraphics[scale=0.35]{./sad-data/dissertation/WeecologyProduction.png}} 
%\subject{} 
\begin{document}

@@

%\begin{frame}
%\tableofcontents
%\end{frame}
\section{Usage Notes}
\begin{frame}
\frametitle{Open science}
\begin{itemize}
\item Code: 
\begin{itemize}
\item github.com/embaldridge
\item github.com/weecology
\end{itemize}
\item Data: figshare.com
\item Twitter: @elitabaldridge
\end{itemize}
\begin{center}
\includegraphics[scale=.40]{./sad-data/dissertation/GitHub-Mark-120px-plus.png}
\includegraphics[scale=.45]{./sad-data/dissertation/figshare_FullColour.jpg} 
\includegraphics[scale=.2]{./sad-data/dissertation/Open_Access.png}  
\includegraphics[scale=.15]{./sad-data/dissertation/osi_logo.png}

\begin{frame}[t]
\begin{huge}
Feel free to:\\
\end{huge}
\includegraphics[scale=.5]{./sad-data/dissertation/use.png}\\
\begin{huge}
Provided that:\\
\end{huge}
\includegraphics[scale=.5]{./sad-data/dissertation/attribution.png}
\end{frame}

\section{Introduction}
\subsection{Commonness & rarity}
\begin{frame}[t]
\frametitle{Commonness \& rarity}
"Who can explain why one species ranges widely, and is very numerous, and why another allied species has a narrow range and is rare?  Yet these relations are of the highest important, for they determine the present welfare and, as I believe, the future success and modification of every inhabitant of this world."

Darwin, 1859.

\note{
Few common species.
Many rare species.
One of the most fundamental and ubiquitous patterns in ecology.}
\end{frame}

\begin{frame}
Species abundance distribution
\begin{itemize}
\item Describes the distribution of commonness \& rarity of species.
\item Exhibits a hollow curve distribution.
\end{itemize}
\includegraphics[scale=.5]{./sad-data/dissertation/generichollowcurves.png}
\note{
Relative abundance distribution (RAD)
Species abundance distribution (SAD)}
\end{frame}

\subsection{What is macroecology?}
\begin{frame}[t]
\frametitle{Macroecology}
\normalsize One approach to studying ecological patterns and processes.\\
\begin{itemize}
\item Data intensive.
\item Large scales.
\begin{itemize}
\item Spatial
\item Temporal
\item Taxonomic
\end{itemize}
\item Search for generality.
\end{itemize}
\end{frame}

\subsection{Signal and Noise}
\begin{frame}[t]
\frametitle{Signal \& Noise}
\begin{center}
\includegraphics[scale=.4]{./sad-data/dissertation/SignalNoise.png}
\end{center}
\end{frame}

\begin{frame}
\frametitle{Macroecology}
\begin{Huge}
\begin{center}

Pattern

\MVArrowDown{}

Process

\MVArrowDown{}

Prediction

\end{center}
\end{Huge}
\end{frame}

@@

\subsection{Best practices}
\begin{frame}[t]
\frametitle{Macroecology}
Best practice recommendations\\
\begin{itemize}
\begin{small}
\item Test patterns with multiple taxonomic groups/ecosystems.  
\item Simultaneous testing of competing models and model predictions with a consistent statistical approach.
\end{small}
\end{itemize}
\end{frame}
biog\subsection{Introduction to Ecoinformatics}
\begin{frame}[t]
\frametitle{The Rules of Ecoinformatics}
\begin{Large}
Garbage in, garbage out.\\
\end{Large}
\begin{itemize}
\item All data are good, not all data are appropriate.
\item Fit the data to the question.
\end{itemize}
\end{frame}

\section{Data}
\subsection{Current Data}
\begin{frame}[t]
\frametitle{Data}
\vspace{-7pt}
\includegraphics[scale=.55]{./sad-data/chapter3/presentation_map.png}
\end{frame}

\begin{frame}[t]{}
\frametitle{Data}
\begin{large}
Major macroecological datasets\\
\end{large}
\begin{itemize}
\item Largely terrestrial
\item Largely North American
\item Many publicly available, some not.
\end{itemize}
~\\
~\\
~\\
~\\
\begin{large}
Lots of data in the literature.\\
\end{large}
\end{frame}

\subsection{Compiled Data}
\begin{frame}{}
\frametitle{Data}
\includegraphics[scale=.35]{../Miscellaneous/Pictures/Phd/CompiledData.png}
\end{frame}

\subsubsection{Data Collection}
\begin{frame}[t]{}
\frametitle{Data}
Inclusion criteria:
\begin{itemize}
\item Quantitative abundances
\item Animals
\item Complete sampling
\item Must not be heavily summarized or processed
\item High degree of taxonomic resolution
\item Observational
\end{itemize}
\end{frame}

@@

\subsubsection{Data Summary}
\begin{frame}{}
\frametitle{Data}
\includegraphics[scale=.27]{./sad-data/dissertation/bioregions.png}
\end{frame}

@@

\subsection{Forms of the distribution}
\begin{frame}[t]
\frametitle{SAD models}
Statistical description\\  
\includegraphics[scale=.2]{./sad-data/dissertation/Preston1948.png}\\

\begin{tiny}
Preston 1962a.\\
\end{tiny}

Process-based\\
\includegraphics[scale=.25]{./sad-data/dissertation/Rosindelletal2011.png}\\
\begin{tiny}
Rosindell et al. 2011.
\end{tiny}
\note{
\begin{itemize}
\item Purely statistical (Lognormal, logseries)
\item Branching process (Zipf)
\item Population dynamics (Neutral theory)
\item Niche partitioning (Geometric)
\item Maximum entropy
\item Feasible set/combinatorics
\end{itemize}}
\end{frame}

@@

\subsubsection{Data}
\begin{frame}[shrink=35]
\frametitle{SAD Comparisons}
\begin{center}
\begin{table}
\begin{tabular}{l|c|l|r}
 Dataset &Dataset code &Availability &Sites\\
\hline
 Gentry's Forest Transects &Gentry &Public &10355\\
 Breeding Bird Survey &BBS &Public &2769\\
 Christmas Bird Count &CBC &Private &1999\\
 Forest Inventory Analysis &FIA	 &Public &220\\
 N. American Butterfly Count &NABA &Private &400\\
 Actinopterygii, compiled &Actinopterygii &Public &161\\
 Reptilia, compiled &Reptilia &Public &138\\
 Mammal Community Database &MCDB &Public &103\\
 Amphibia, compiled &Amphibia &Public &43\\
 Arachnida, compiled &Arachnida &Public &25\\
 Coleoptera, compiled &Coleoptera &Public &5\\
\end{tabular}
\caption{Datasets used for species-abundance distribution comparisons. Datasets marked as Private obtained through data requests to the providers with Memorandums of Understanding.}
\end{table}
\end{center}
\end{frame}

\begin{frame}{}
\frametitle{Data}
\includegraphics[scale=.4]{../Miscellaneous/Pictures/Phd/MapCreatures.png}
\end{frame}

@@

\begin{frame}[t]
\frametitle{SAD Comparisons}
Identifying pattern generating mechanisms:
\begin{itemize}
\item Compare predictions of different models using multiple macroecological patterns simultaneously.\\
\item Examine scale dependence of pattern.
\end{itemize}
~\\
~\\
However, identification of mechanism may not be necessary for prediction.
\end{frame}

\section{Neutral analysis}
\begin{frame}[t]
\frametitle{Neutral theory}
~\\
Many formulations, but:
~\\
\begin{itemize}
\item Species \& individuals ecologically \& demographically equivalent.\\
\end{itemize}
\includegraphics[scale=.35]{./sad-data/dissertation/Rosindelletal2011.png}\\
\begin{tiny}
Rosindell et al. 2011.
\end{tiny}
\note{
\begin{itemize}
\item Species \& individuals ecologically \& demographically equivalent.\\
\item Stochastic variation in birth, death, immigration, \& speciation results in species abundance differences.
\end{itemize}}
\end{frame}

\begin{frame}
\frametitle{Neutral Analysis}
Early tests of neutral theory compared the fit of empirical species abundance distributions to the neutral prediction.\\
~\\
Later tests suggested species abundance comparisons were insufficient for a rigorous test of neutrality.\\
~\\
\begin{Large}
\emph{However...}
\end{Large}
\end{frame}

@@

@@

\subsection{Results}
\begin{frame}{}
\frametitle{Neutral Analysis}
\includegraphics[scale=.45]{./sad-data/dissertation/distabclasses_vs_lognormwgt.png}
\end{frame}

\begin{frame}{}
\frametitle{Neutral Analysis}
\includegraphics[scale=.40]{./sad-data/dissertation/avgvals_by_dataset.png}
\end{frame}

@@

\begin{frame}[t]{}
\frametitle{Conclusions}\
~\\ 
\begin{large}
Predictive macroecology
\end{large}
~\\ 
\begin{itemize}
\item Traditional approach is pattern to process to prediction.
\item May be possible to generate robust ecological predictions from general patterns.
\item Process and prediction may be two separate research goals.
\end{itemize} 
\end{frame}

@@

\subsection{Accessibility}
\begin{frame}[t]{}
\frametitle{Accessibility}
\begin{center}
This dissertation brought to you by:
\end{center}
\begin{Large}
Disability accommodations\\
\end{Large}
~\\
\begin{itemize}
\item Remote access \& participation.\\
\item Computational tools \& tricks.
\begin{itemize}
\item Version control (GitHub).
\item Publicly available data.
\item Programming skills (data manipulation \& analysis).
\end{itemize}
\end{itemize}
\end{frame}

>>>>

\documentclass[14pt]{beamer}
\usetheme{Pittsburgh}
\usepackage{xcolor}
\definecolor{USUBlue}{RGB}{26,57,89}
\usecolortheme[named=USUBlue]{structure}
\setbeamercolor{normal text}{fg=USUBlue}
\usepackage[utf8]{inputenc}
\usepackage{mathptmx}
\usepackage{tgbonum}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{marvosym}
\usefonttheme{structuresmallcapsserif}
\usefonttheme{serif}
\setbeamercolor{author}{fg=USUBlue}
\setbeamerfont{author}{size=\small}
\setbeamerfont{frametitle}{size=\large}
\setbeamertemplate{enumerate items}[default]
\author{Elita Baldridge\\ @elitabaldridge}
\title[17pt]{A data-intensive assessment of the species-abundance distribution.}
\setbeamertemplate{navigation symbols}{}
\date{}
%\setbeamercovered{transparent} 
%\logo{\includegraphics[scale=.03]{../Miscellaneous/Pictures/ecology_center_horizontal.jpg}\includegraphics[scale=0.07]{../Miscellaneous/Pictures/Weecology.png}} 
\institute{\includegraphics[scale=.07]{../Miscellaneous/Pictures/ecology_center_horizontal.jpg}\includegraphics[scale=0.35]{./sad-data/dissertation/WeecologyProduction.png}} 
%\subject{} 
\begin{document}

@@

%\begin{frame}
%\tableofcontents
%\end{frame}
\section{Usage Notes}
\begin{frame}[t]
\begin{huge}
Feel free to:\\
\end{huge}
\includegraphics[scale=.5]{./sad-data/dissertation/use.png}\\
\begin{huge}
Provided that:\\
\end{huge}
\includegraphics[scale=.5]{./sad-data/dissertation/attribution.png}
\end{frame}

\begin{frame}
\frametitle{Open science}
\begin{itemize}
\item Code: 
\begin{itemize}
\item github.com/embaldridge
\item github.com/weecology
\end{itemize}
\item Data: figshare.com
\end{itemize}
\begin{center}
\includegraphics[scale=.40]{./sad-data/dissertation/GitHub-Mark-120px-plus.png}
\includegraphics[scale=.45]{./sad-data/dissertation/figshare_FullColour.jpg} 
\includegraphics[scale=.2]{./sad-data/dissertation/Open_Access.png}  
\includegraphics[scale=.15]{./sad-data/dissertation/osi_logo.png}
\end{center}
\end{frame}

\section{Introduction}
\subsection{What is macroecology?}
\begin{frame}[t]
\frametitle{Macroecology}
\normalsize Search for general patterns \& processes.\\
\begin{center}
\includegraphics[scale=.35]{./sad-data/dissertation/PatternProcess_example.png}
\end{center}
\note{
\begin{itemize}
\item Data intensive.
\item Large scales.
\begin{itemize}
\item Spatial
\item Temporal
\item Taxonomic
\end{itemize}
\item Search for general patterns \& processes.
\end{itemize}}
\end{frame}

\begin{frame}
Species abundance distribution
\note{
Few common species.
Many rare species.
One of the most fundamental and ubiquitous patterns in ecology.}
\begin{itemize}
\item Describes the distribution of commonness \& rarity of species.
\item Exhibits a hollow curve distribution.
\end{itemize}
\includegraphics[scale=.5]{./sad-data/dissertation/generichollowcurves.png}\\
\begin{itemize}
\item Many potential pattern-generating mechanisms.
\end{itemize}
\note{
Relative abundance distribution (RAD)
Species abundance distribution (SAD)}
\end{frame}

\begin{frame}
\frametitle{Pattern \& Process;\\ Signal \& Noise}
\begin{center}
\includegraphics[scale=.4]{./sad-data/dissertation/SignalNoise.png}
\end{center}
\note{
\begin{Huge}
\begin{center}
Pattern\\ 
\MVArrowDown{}\\  
Process\\  
\MVArrowDown{}\\
Prediction
\end{center}
\end{Huge}}
\end{frame}

\subsection{Commonness & rarity}
%\begin{frame}[t]
%\frametitle{Commonness \& rarity}
%"Who can explain why one species ranges widely, and is very numerous, and why another %allied species has a narrow range and is rare?  Yet these relations are of the highest %important, for they determine the present welfare and, as I believe, the future success %and modification of every inhabitant of this world."\\ 
%~\\ 
%Darwin, 1859.\\
%\end{frame}

@@

\subsection{Best practices}
\begin{frame}[t]
\frametitle{Macroecology}
Best practice recommendations\\
\begin{itemize}
\item Simultaneous testing of competing models and model predictions with a consistent statistical approach.
\item Test patterns with multiple taxonomic groups/ecosystems.  
\end{itemize}
\end{frame}

\section{Data}
\subsection{Current Data}
\begin{frame}[t]
\frametitle{Data}
\vspace{-7pt}
\includegraphics[scale=.55]{./sad-data/chapter3/presentation_map.png}
\note{
\begin{large}
Major macroecological datasets\\
\end{large}
\begin{itemize}
\item Largely terrestrial
\item Largely North American
\item Many publicly available, some not.
\end{itemize}
~\\
~\\
~\\
~\\
\begin{large}
Lots of data in the literature.\\
\end{large}}
\end{frame}

\subsection{Compiled Data}
\subsubsection{Introduction to Ecoinformatics}
\begin{frame}[t]
\frametitle{The Rules of Ecoinformatics}
\begin{Large}
Garbage in, garbage out.\\
\end{Large}
\begin{itemize}
\item All data are good, not all data are appropriate.
\item Fit the data to the question.
\end{itemize}
\end{frame}
\subsubsection{Data Collection}
\begin{frame}[t]{}
\frametitle{Data}
Inclusion criteria:
\begin{itemize}
\item Quantitative abundances
\item Animals
\item Complete sampling
\item Must not be heavily summarized or processed
\item High degree of taxonomic resolution
\item Observational
\end{itemize}
\end{frame}

@@

\subsubsection{Data Summary}
\begin{frame}{}
\frametitle{Data}
\includegraphics[scale=.35]{../Miscellaneous/Pictures/Phd/CompiledData.png}
\end{frame}

\begin{frame}{}
\frametitle{Data}
\includegraphics[scale=.27]{./sad-data/dissertation/bioregions.png}
\end{frame}

@@

\subsection{Forms of the distribution}
\begin{frame}[t]
\frametitle{SAD models}
Statistical description\\
\begin{center}  
\includegraphics[scale=.2]{./sad-data/dissertation/Preston1948.png}
\begin{tiny}
Preston 1962a.\\
\end{tiny}
\end{center}
Process-based\\
\begin{center}
\includegraphics[scale=.25]{./sad-data/dissertation/Rosindelletal2011.png}
\begin{tiny}
Rosindell et al. 2011.\\
\end{tiny}
\end{center}
\note{
\begin{itemize}
\item Purely statistical (Lognormal, logseries)
\item Branching process (Zipf)
\item Population dynamics (Neutral theory)
\item Niche partitioning (Geometric)
\item Maximum entropy
\item Feasible set/combinatorics
\end{itemize}}
\end{frame}

@@

\subsubsection{Data}
\begin{frame}{}
\frametitle{Data}
\includegraphics[scale=.4]{../Miscellaneous/Pictures/Phd/MapCreatures.png}
\end{frame}

\begin{frame}[shrink=35]
\frametitle{SAD Comparisons}
\begin{center}
\begin{table}
\begin{tabular}{l|c|l|r}
 Dataset &Dataset code &Availability &Sites\\
\hline
 Gentry's Forest Transects &Gentry &Public &10355\\
 Breeding Bird Survey &BBS &Public &2769\\
 Christmas Bird Count &CBC &Private &1999\\
 Forest Inventory Analysis &FIA	 &Public &220\\
 N. American Butterfly Count &NABA &Private &400\\
 Actinopterygii, compiled &Actinopterygii &Public &161\\
 Reptilia, compiled &Reptilia &Public &138\\
 Mammal Community Database &MCDB &Public &103\\
 Amphibia, compiled &Amphibia &Public &43\\
 Arachnida, compiled &Arachnida &Public &25\\
 Coleoptera, compiled &Coleoptera &Public &5\\
\end{tabular}
\caption{Datasets used for species-abundance distribution comparisons. Datasets marked as Private obtained through data requests to the providers with Memorandums of Understanding.}
\end{table}
\end{center}
\end{frame}

@@

\begin{frame}[t]
\frametitle{SAD Comparisons}
Identifying pattern generating mechanisms:
\begin{itemize}
\item Compare predictions of different models using multiple macroecological patterns simultaneously.\\
\item Examine scale dependence of pattern.
\item More general approach to process.
\end{itemize}
\end{frame}

\section{Neutral analysis}
\begin{frame}
\frametitle{Neutral Analysis}
Early tests of neutral theory compared the fit of empirical species abundance distributions to the neutral prediction.\\
~\\
Later tests suggested species abundance comparisons were insufficient for a rigorous test of neutrality.\\
~\\
\begin{Large}
\emph{However...}
\end{Large}
\end{frame}

@@

\begin{frame}[t]
\frametitle{Neutral theory}
~\\
Many formulations, but:
~\\
\begin{itemize}
\item Species \& individuals ecologically \& demographically equivalent.\\
\end{itemize}
\includegraphics[scale=.35]{./sad-data/dissertation/Rosindelletal2011.png}\\
\begin{tiny}
Rosindell et al. 2011.
\end{tiny}
\note{
\begin{itemize}
\item Species \& individuals ecologically \& demographically equivalent.\\
\item Stochastic variation in birth, death, immigration, \& speciation results in species abundance differences.
\end{itemize}}
\end{frame}

@@

\subsection{Results}
\begin{frame}{}
\frametitle{Neutral Analysis}
\includegraphics[scale=.40]{./sad-data/dissertation/avgvals_by_dataset.png}
\end{frame}

\begin{frame}{}
\frametitle{Neutral Analysis}
\includegraphics[scale=.45]{./sad-data/dissertation/distabclasses_vs_lognormwgt.png}
\end{frame}

@@

%\begin{frame}[t]{}
%\frametitle{Conclusions}\
%~\\ 
%\begin{large}
%Predictive macroecology
%\end{large}
%~\\ 
%\begin{itemize}
%\item Traditional approach is pattern to process to prediction.
%\item May be possible to generate robust ecological predictions from general patterns.
%\item Process and prediction may be two separate research goals.
%\end{itemize} 
%\end{frame}

@@

\subsection{Accessibility}
\begin{frame}[t]{}
\frametitle{Accessibility}
\begin{center}
This dissertation brought to you by:
\end{center}
\begin{Large}
Disability accommodations\\
\end{Large}
~\\
\begin{itemize}
\item Remote access \& participation.\\
\item Accessibility information
\note{
\item Computational tools \& tricks.
\begin{itemize}
\item Version control (GitHub).
\item Publicly available data.
\item Programming skills (data manipulation \& analysis).
\end{itemize}}
\end{itemize}
\end{frame}
